Military mechanization as a complex adaptive system

### 2018-07-20 paper2/paper2.tex
so close

MORE ON VERIFICATION

The main model output of interest is data on military mechanization. It is provided for
every agent and every time-step. These agent-level data also can be aggregated
into system-level measures of military mechanization. Per the research design,
the intent is to compare the model output to what actually occurred. If the
model produces military mechanization levels much higher than reality, it means
that Sechser and Saunder's findings do not stand up to agentization and there is
some missing ``brake'' on military mechanization. Note that this research
strategy has validation (comparison of model output to empirical data) at its
core and so no separate validation is needed. Verification that the model is
doing what it is intended was accomplished through manual inspection.

@@

For some agents, however, mechanization levels are extremely high. 
Mechanization score is a ratio, and so values over 1 do not make sense, 
yet some countries end the simulation with mechanization scores over 1. Note
that one could easily right-bound values at 1, but this was not done in order to
see what values would be organically converged to.

@@

BEEF UP THIS SECTION

Two conclusions should be drawn. One, the model results in much higher
system-level mechanization than occurred in reality. Two, this higher
mechanization is not evenly distributed across agents; rather, a small number of
agents with many enemies and that all border each other become outliers and
hence drive the global average far upward. The explanation for both phenomena
lies in a positive feedback loop. Enemies and neighbors adjust their
mechanization scores in response to each other; hence, a geographic cluster of
enemies (e.g., the Middle East or Cold War-era Europe) will rapidly increase their
mechanization.

Three, because this model is a
faithful representation of the factors that \citet{sechser2010army} claim are
most important for mechanization, that model output fails to match reality
indicates that their claimed model is incomplete. There must be some ``brake''
on mechanization; put differently, some countervailing factor(s) must mitigate the
effect of the positive feedback loop identified above. This model cannot
conclusively identify what that ``brake'' is. The most plausible candidates are
factors examined by \citet{sechser2010army} and deemed statistically
insignificant such as regime type and economy, and factors not examined by
\citet{sechser2010army} at all such as culture.

>>>>

The main model output of interest is data on military mechanization. It is provided for
every agent and every time-step. These agent-level data also can be aggregated
into system-level measures of military mechanization.

Verification--—i.e., making sure that the model does what it is intended to
do---is a relative activity, not absolute. It is intended to increase confidence in the correctness
of the model, not to prove that it is perfect \citep{robinson1997simulation}.
During model development, the model was configured for verbose run-time output. Manual
inspection of this output confirmed that the model was working as intended. For
example, output of each actor's \textit{Enemies} at each timestep allowed
confirmation that the actor was updating its perception in response to scripted
international conflict in line with the historical record.

Validation---i.e., assessing the correspondence between the model and empirical
data---is central to the research strategy. The National Mechanization Index
(NMI) compiled by \citet{sechser2010army} provides empirical data on
mechanization at the country-year level between 1979 and 2001. In the following
\textbf{Results} section, model output is compared to this data in
order to test the hypothesis that the orthodox view of mechanization does
not stand up to agentization. Because the
model and empirical data are at the country-year level, validation can take
place at both the micro-level (i.e., individual state-agents) and at the
macro-level (i.e., the aggregate mechanization levels of the whole international
system); and, validation is longitudinal rather than static. 
Thus, this paper pursues Level 3 validation in the taxonomy of
\citet{axtell1994understanding}: ``cross-sectional and longitudinal analysis of
the agent population.''

@@

For some agents, however, mechanization levels are extremely high. 
Mechanization score is a ratio, and so values over 1 do not make sense in
context, 
yet some countries end the simulation with mechanization scores over 1. Note
that one could easily right-bound values at 1, but this was not done in order to
see what values would be organically converged to.

@@

BEEF THIS UP AND YOU'RE DONE

The model results in much higher
system-level mechanization than occurred in reality. This higher
mechanization is not evenly distributed across agents. Rather, a small number of
agents with many enemies and that all border each other become outliers and
hence drive the global average far upward. The explanation for both phenomena
lies in a positive feedback loop. Enemies and neighbors adjust their
mechanization scores in response to each other; hence, a geographic cluster of
enemies (e.g., the Middle East or Cold War-era Europe) will rapidly increase their
mechanization.

Because this model is a faithful representation of the factors that \citet{sechser2010army} claim are
most important for mechanization, that model output fails to match reality
indicates that their claimed model is incomplete. There must be some ``brake''
on mechanization; put differently, some countervailing factor(s) must mitigate the
effect of the positive feedback loop identified above. This model cannot
conclusively identify what that ``brake'' is. The most plausible candidates are
factors examined by \citet{sechser2010army} and deemed statistically
insignificant such as regime type and economy, and factors not examined by
\citet{sechser2010army} at all such as culture.
